Optimal updating factor in adaptive flat-distribution-sampling simulations

### 2017-06-17 doc/response.tex

\documentclass{article}
\usepackage{amsmath}
\begin{document}

The old definition of $\rho_i$ and $p_i$ are assumed to be same now
to avoid unnecessary complexity, preserving the old $\rho_i$ only.
The old $p^*_i$ is replaced by the new $p_i$.
The old $V_i$ is replaced by $u_i$.
The old $x_i$ is replaced by $v_{*i}$.
The old $y_k$ is replaced by $\tilde v_k$.
The old $\theta_k$ is replaced by $\tilde f_k$.
The old $\zeta_i/p_i$ is replaced by $\zeta_i$.
The old $C$ is replaced by $C_M$.
The old degeneracy $g$ is replaced by $2/g$ for the multiple of period.
%
We have defined $f_i \equiv h_i/\rho_i$.
We changed $Q \to \bar q$ to imply the meaning of complement.

\end{document}

### 2017-06-17 doc/invt.tex
invt.tex: add back sec:seplambda

As in the single-bin case,
by varying the schedule, $\alpha(t)$, under a fixed value of
$q(T) = \int_0^T \alpha(t) \, dt$,
we can focus on the asymptotic error,
$\Err_A(T)$.
%
The value of $q(T)$ will be adjusted later
to minimize the total error in Sec. \ref{sec:optinitalpha}).
%
We can rewrite Eq. \eqref{eq:error_asym} much like an action
of a particle whose position is given by $q(t)$:
%
\begin{equation}
  \Err_A(T)
  =
  \int_0^T
    {\mathcal L} \bigl[ q(t)\bigr]
    \, dt
  ,
%\notag
\label{eq:error_asym_Lagrangian}
\end{equation}
%
where the Lagrangian is
%
\begin{align}
  {\mathcal L} \bigl[ q(t) \bigr]
  &=
  \sum_{ k = 1 }^{n - 1}
    \Gamma_k \, {\dot y}_k^2\bigl[ q(t) \bigr]
  %\notag
  %\\
  %&=
  =
  %\sum_{ k = 1 }^{n - 1}
  %  \Gamma_k \, \lambda_k^2 \, y_k^2 \bigl[ q(t) \bigr]
  %\; \dot q^2( t )
  %\notag
  %\\
  %&=
  M^2\bigl(q(T) - q(t) \bigr)
  \; \dot q^2( t )
  .
\notag
\end{align}
%
where we have
used Eq. \eqref{eq:uk_def} and
defined the (square-root) mass function as
%
\begin{equation}
  M(q')
  \equiv
  \sqrt{
    \textstyle\sum_{ k = 1 }^{n - 1}
    \Gamma_k \, \lambda_k^2 \, e^{-2 \, \lambda_k \, q'}
  }
  .
  \notag
  %\label{eq:mass_func}
\end{equation}
%
In this mechanical analogy,
the schedule, $\dot q(t) = \alpha(t)$,
corresponds to the velocity of the particle,
%
and the above Lagrangian, containing only the kinetic energy,
describes a free particle
with a position-dependent mass.
%
Besides, the Hamiltonian
coincides with the Lagrangian:
%
\begin{equation}
  \mathcal H
  =
  \frac{ \partial \mathcal L }
       { \partial \dot q     }
  \, \dot q
  -
  \mathcal L
  =
  2 \, \mathcal L
  - \mathcal L
  =
  \mathcal L
  .
  %=
  %\mathrm{const.}
  \notag
  %\label{eq:error_asym_Hamiltonian}
\end{equation}
%
Since the Lagrangian
does not explicitly depend on time, $t$,
the Hamiltonian is conserved,
which means that the asymptotic error grows
linearly with time at a rate of $\mathcal L$,
and we may set
%
\begin{equation}
  \sqrt{ \mathcal H }
  =
  \sqrt{ \mathcal L }
  =
  M\bigl( q(T) - q(t) \bigr)
  \;
  \dot q(t)
  =
  \frac{C}{T}
  ,
  \label{eq:Lagrangian_const}
\end{equation}
%
for some positive $C$.
%
This serves as a generalization of Eq. \eqref{eq:dydt_const}.
%
By integrating this equation of motion, we get
%
\begin{equation}
  \int_{ q(T) - q(t) }^{ q(T) }
    M(q')
    \;
    d q'
  =
  C \, \frac t T
  ,
  \label{eq:q_opt}
\end{equation}
%
and $C$ can be determined from
the $t = T$ case as
%
\begin{equation}
  C =
  \int_{ 0 }^{ q(T) }
    M( q' )
    \;
    d q'
  .
  \label{eq:mint}
\end{equation}
%
Thus, we have an implicit equation for $q(t)$,
and hence the optimal schedule,
$\alpha(t) = \dot q(t)$.
%
The minimal asymptotic error is
%
\begin{equation}
  \Err_A(T)
  =
  \mathcal L \, T
  =
  \frac { C^2 } { T }
  =
  \frac 1 T
  \left(
    \int_0^{ q(T) } M(q') \, d q'
  \right)^2
  .
\label{eq:error_asym2}
\end{equation}

@@

We can characterize the optimal schedule
geometrically as follows.
%
First, we define
a normalized mass distribution as
%
\begin{equation}
  m(\bar q)
  =
  \frac{
    M(\bar q)
  }
  {
    \int_0^{ q(T) } M(q') \, d q'
  }
  =
  \frac{
    M(\bar q)
  }
  {
    C
  }
  ,
\notag
%\label{eq:mass_distr}
\end{equation}
%
such that
$\int_0^{q(T)} m(\bar q) \, d\bar q = 1$.
%
Then from Eqs. \eqref{eq:Lagrangian_const} and \eqref{eq:q_opt},
we find that for $\bar q(t) \equiv q(T) - q(t)$,
%
\begin{align}
  m\left( \bar q(t) \right)
  &=
  \frac{ 1 }
       { T \, \alpha(t) }
  ,
  \label{eq:mQ_invTa}
  \\
  \int_{\bar q(t)}^{ q(T) }
    m(q') \, dq'
  &=
  \frac t T
  ,
  \label{eq:intmQ_tT}
\end{align}
%
Thus, each point on the schedule,
$\bigl(t, \alpha(t)\bigr)$,
can be mapped to a point,
$\Bigl(\bar q(t), m\bigl(\bar q(t)\bigr)\Bigr)$,
on the mass distribution,
such that the ordinate of the latter curve
is $1/(T\alpha)$,
and the area under the curve in the domain, $[\bar q(t), q(T)]$,
is equal to $t/T$,
as shown in Fig. \ref{fig:massq}.

@@

We may consider the schedule, $\alpha(t)$,
as a variable transformation of
the mass distribution, $m(\bar q)$.
%much like the Legendre transform.
%
If we define
$\bar q^* \equiv \bar q/q(T)$
and $t^* \equiv t/T$
such that both are limited in $[0, 1]$,
as well as
\begin{align*}
  m^*       &\equiv q(T) \, m(\bar q) = -\frac{ dt^* } { d\bar q^* },
  \\
  \alpha^*  &\equiv \frac{T}{q(T)} \, \alpha(t) = -\frac{ d\bar q^* } { dt^* },
\end{align*}
then it is clear that
both $m^*(\bar q^*)$ and $\alpha^*(t^*)$
can describe the differential relation
between $\bar q^*$ and $t^*$,
but with the natural variables
being $\bar q^*$ and $t^*$, respectively.
%
Further, the transformation is an involution:
if we apply the same transform on $\alpha^*(t^*)$
we would recover $m^*(\bar q^*)$.
%
A few examples are shown in Table \ref{tab:m_and_a}.
%
For example, in the single-bin scheme,
the mass distribution is a single exponential decay,
and we would have the inverse-time schedule.
%
On the other hand,
if the mass distribution is inversely proportional to $\bar q$,
which is approximately true for the Gaussian updating scheme
at large $\bar q$,
then we would have an exponentially decaying schedule.

@@

We now determine the optimal value of $q(T)$
that minimizes the total error defined in Eq. \eqref{eq:error_tot}.
%
We will consider simulations in which
the system is initially equilibrated at a
constant updating magnitude, $a_0$.
%
From Eqs. \eqref{eq:error_res}, \eqref{eq:xt2_eql1},
and \eqref{eq:error_asym2},
we find that
the optimal $q(T)$ satisfies
%
\begin{align}
  \mathcal Y\bigl( q(T) \bigr)
  &\equiv
  \frac{ M\bigl( q(T) \bigr) } { T } \int_0^{q(T)} M(q') \, dq'
  -\frac{a_0}{2} \, M^2\bigl( q(T) \bigr)
  \notag \\
  &
  -\sum_{k = 1}^{n-1}
  \lambda_k \, \epsilon_k \, e^{-2\lambda_k \, q(T)}
  =0
  .
\label{eq:opt_qT}
\end{align}
%
%\begin{align}
%  m\bigl( q(T) \bigr)
%  +
%  \frac{2}{a_0 \, C \, M\bigl( q(T) \bigr)}
%  \sum_{k = 1}^{n-1}
%  \lambda_k \, \tilde u_k^2 \,
%  e^{-2 \, \lambda_k \, [a_0 \, T_0 + q(T)] }
%  \\=
%  \frac{1} { T \, (a_0 / 2) }
%  .
%\label{eq:opt_qT}
%\end{align}
%
\note{Let $q_T \equiv q(T)$,
$$
\begin{aligned}
  \frac{
    \partial \Err_R(T)
  }
  {
    \partial q_T
  }
  &=
  -a_0 \, M^2(q_T)
  -2
  \sum_{k=1}^{n-1} \lambda_k \,
  \tilde u_k^2 e^{-2 \, \lambda_k \, (a_0 \, T_0 + q_T)}
  ,
  \\
  \frac{
    \partial \Err_A(T)
  }
  {
    \partial q_T
  }
  &=
  \frac 2 T \,
  M(q_T) \,
  \int_0^{ q_T } M(q) \, dq
  =
  \frac{ 2 \, C } { T } \, M(q_T)
  .
\end{aligned}
$$
Then solve $\partial (\Err_R + \Err_A) / \partial Q_c = 0$.
}
%
We can rewrite the condition in terms of
the initial updating magnitude by Eq. \eqref{eq:mQ_invTa},
\begin{align}
  \alpha(0)
  =
  \frac{1}{T \, m\bigl( q(T) \bigr) }
  =
  \frac{ a_0 } { 2 }
  +
  \sum_{k = 1}^{n-1}
  \frac{
    \lambda_k \, \epsilon_k \, e^{-2\lambda_k \, q(T)}
  }{M^2\bigl( q(T) \bigr)}
  .
  \label{eq:opt_alpha0}
\end{align}
%
If the initial bias, $\epsilon_k$, is negligible,
which is typically true if no eigenvalue is close to $0$,
then
%
\begin{equation}
  \alpha( t = 0 )
  \approx
  \frac{ a_0 }
       { 2 }
  ,
\label{eq:half_alpha0}
\end{equation}
%
i.e. the optimal initial updating magnitude
is half of the equilibrium value.\footnote{This factor, $1/2$,
happens to coincide with the
recommended reduction factor of the updating magnitude
at stage transitions
in the original WL algorithm\cite{
wang2001, *wang2001pre}.
}

@@

The transition matrix method usually only for
simple sampling processes without major barrier
along the collective variable, $z$.
%
We may nonetheless estimate the magnitude of $\Gamma_k$
in a few special cases.
%
If sampling is perfect, $\Lambda_k = \delta_{k, 0}$,
then $\Gamma_k = 1$ for $k > 0$.
%
For a local sampling process,
if we borrow \eqref{eq:lambda_Gaussian}
for $\Lambda_k$,
with $\sigma_z$ interpreted as the average move size along $z$,
then
\begin{align*}
\Gamma_k = \coth\left[
  \left(
    \frac{ \pi \, k \, \sigma_z } { g \, \Delta z}
  \right)^2
  \right].
\end{align*}
For small $k$,
we have $\Gamma_k \propto \Delta z^2/(\sigma_z \, k)^2$,\cite{bussi2006}
which means that a doubled sampling range, $\Delta z$,
would make the histogram flattening
at least four times as difficult,
i.e. FDS methods are most effective
for short ranges\cite{wang2001, *wang2001pre}.
%
Using finer bins would, however, not change this value.
%
For a modestly large $k$,
$\Gamma_k \approx 1$.
and by Eq. \eqref{eq:xt2_eql},
we would expect a roughly flat profile
for the single-bin updating scheme,
cf. Fig. \ref{fig:lj_xerr}.

>>>>

As in the single-bin case,
by varying the schedule, $\alpha(t)$, under a fixed value of
$q(T) = \int_0^T \alpha(t) \, dt$,
we can focus on the asymptotic error,
$\Err_A(T)$.
%
The value of $q(T)$ will be adjusted later
to minimize the total error in Sec. \ref{sec:optinitalpha}).
%
We can rewrite Eq. \eqref{eq:error_asym} much like an action
of a particle whose position is given by $q(t)$:
%
\begin{equation}
  \Err_A(T)
  =
  \int_0^T
    {\mathcal L} \bigl[ q(t)\bigr]
    \, dt
  ,
%\notag
\label{eq:error_asym_Lagrangian}
\end{equation}
%
where the Lagrangian is
%
\begin{align}
  {\mathcal L} \bigl[ q(t) \bigr]
  &=
  \sum_{ k = 1 }^{n - 1}
    \Gamma_k \, {\dot y}_k^2\bigl[ q(t) \bigr]
  %\notag
  %\\
  %&=
  =
  %\sum_{ k = 1 }^{n - 1}
  %  \Gamma_k \, \lambda_k^2 \, y_k^2 \bigl[ q(t) \bigr]
  %\; \dot q^2( t )
  %\notag
  %\\
  %&=
  M^2\bigl(q(T) - q(t) \bigr)
  \; \dot q^2( t )
  .
\notag
\end{align}
%
where we have
used Eq. \eqref{eq:uk_def} and
defined the (square-root) mass function as
%
\begin{equation}
  M(\bar q)
  \equiv
  \sqrt{
    \textstyle\sum_{ k = 1 }^{n - 1}
    \Gamma_k \, \lambda_k^2 \, e^{-2 \, \lambda_k \, \bar q}
  }
  .
  \label{eq:mass_func}
\end{equation}
%
In this mechanical analogy,
the schedule, $\dot q(t) = \alpha(t)$,
corresponds to the velocity of the particle,
%
and the above Lagrangian, containing only the kinetic energy,
describes a free particle
with a position-dependent mass.
%
Besides, the Hamiltonian
coincides with the Lagrangian:
%
\begin{equation}
  \mathcal H
  =
  \frac{ \partial \mathcal L }
       { \partial \dot q     }
  \, \dot q
  -
  \mathcal L
  =
  2 \, \mathcal L
  - \mathcal L
  =
  \mathcal L
  .
  %=
  %\mathrm{const.}
  \notag
  %\label{eq:error_asym_Hamiltonian}
\end{equation}
%
Since the Lagrangian
does not explicitly depend on time, $t$,
the Hamiltonian is conserved,
which means that the asymptotic error grows
linearly with time at a rate of $\mathcal L$,
and we may set
%
\begin{equation}
  \sqrt{ \mathcal H }
  =
  \sqrt{ \mathcal L }
  =
  M\bigl( q(T) - q(t) \bigr)
  \;
  \dot q(t)
  =
  \frac{C_M}{T}
  ,
  \label{eq:Lagrangian_const}
\end{equation}
%
for some positive $C_M$.
%
This serves as a generalization of Eq. \eqref{eq:dydt_const}.
%
By integrating this equation of motion, we get
%
\begin{equation}
  \int_{ q(T) - q(t) }^{ q(T) }
    M(\bar q)
    \;
    d \bar q
  =
  C_M \, \frac t T
  ,
  \label{eq:q_opt}
\end{equation}
%
and $C_M$ can be determined from
the $t = T$ case as
%
\begin{equation}
  C_M =
  \int_{ 0 }^{ q(T) }
    M( \bar q )
    \;
    d \bar q
  .
  \label{eq:mint}
\end{equation}
%
Thus, we have an implicit equation for $q(t)$,
and hence the optimal schedule,
$\alpha(t) = \dot q(t)$.
%
The minimal asymptotic error is
%
\begin{equation}
  \Err_A(T)
  =
  \mathcal L \, T
  =
  \frac { C_M^2 } { T }
  =
  \frac 1 T
  \left(
    \int_0^{ q(T) } M(\bar q) \, d \bar q
  \right)^2
  .
\label{eq:error_asym2}
\end{equation}

@@

We can characterize the optimal schedule
geometrically as follows.
%
First, we define
a normalized mass distribution as
%
\begin{equation}
  m(\bar q)
  =
  \frac{
    M(\bar q)
  }
  {
    \int_0^{ q(T) } M(q') \, d q'
  }
  =
  \frac{
    M(\bar q)
  }
  {
    C_M
  }
  ,
\notag
%\label{eq:mass_distr}
\end{equation}
%
such that
$\int_0^{q(T)} m(\bar q) \, d\bar q = 1$.
%
Then from Eqs. \eqref{eq:Lagrangian_const} and \eqref{eq:q_opt},
we find that for $\bar q(t) \equiv q(T) - q(t)$,
%
\begin{align}
  m\left( \bar q(t) \right)
  &=
  \frac{ 1 }
       { T \, \alpha(t) }
  ,
  \label{eq:mQ_invTa}
  \\
  \int_{\bar q(t)}^{ q(T) }
    m(q') \, d q'
  &=
  \frac t T
  ,
  \label{eq:intmQ_tT}
\end{align}
%
Thus, each point on the schedule,
$\bigl(t, \alpha(t)\bigr)$,
can be mapped to a point,
$\Bigl(\bar q(t), m\bigl(\bar q(t)\bigr)\Bigr)$,
on the mass distribution,
such that the ordinate of the latter curve
is $1/(T\alpha)$,
and the area under the curve in the domain, $[\bar q(t), q(T)]$,
is equal to $t/T$,
as shown in Fig. \ref{fig:massq}.

@@

We may consider the schedule, $\alpha(t)$,
as a variable transformation of
the mass distribution, $m(\bar q)$.
%much like the Legendre transform.
%
If we define
$\bar q^* \equiv \bar q/q(T)$
and $t^* \equiv t/T$
such that both are limited in $[0, 1]$,
as well as
\begin{align*}
  m^*       &\equiv q(T) \, m(\bar q) = -\frac{ dt^* } { d\bar q^* },
  \\
  \alpha^*  &\equiv \frac{T}{q(T)} \, \alpha(t) = -\frac{ d\bar q^* } { dt^* },
\end{align*}
then it is clear that
both $m^*(\bar q^*)$ and $\alpha^*(t^*)$
can describe the differential relation
between $\bar q^*$ and $t^*$,
but with the natural variables
being $\bar q^*$ and $t^*$, respectively.
%
Further, the transformation is an involution:
if we apply the same transform on $\alpha^*(t^*)$
we would recover $m^*(\bar q^*)$.
%
A few examples are shown in Table \ref{tab:m_and_a}.
%
For example, in the single-bin scheme,
the mass distribution is a single exponential decay,
and we would have the inverse-time schedule.
%
On the other hand,
if the mass distribution is inversely proportional to $\bar q$,
which is approximately true for the Gaussian updating scheme
at large $\bar q$ (cf. Appendix \ref{sec:seplambda}),
then we would have an exponentially decaying schedule.

@@

We now determine the optimal value of $q(T)$
that minimizes the total error defined in Eq. \eqref{eq:error_tot}.
%
We will consider simulations in which
the system is initially equilibrated at a
constant updating magnitude, $a_0$.
%
From Eqs. \eqref{eq:error_res}, \eqref{eq:xt2_eql1},
and \eqref{eq:error_asym2},
we find that
the optimal $q(T)$ satisfies
%
\begin{align}
  \mathcal Y\bigl( q(T) \bigr)
  &\equiv
  \frac{ M\bigl( q(T) \bigr) } { T }
    \int_0^{q(T)} M(q') \, d q'
  -\frac{a_0}{2} \, M^2\bigl( q(T) \bigr)
  \notag \\
  &
  -\sum_{k = 1}^{n-1}
  \lambda_k \, \epsilon_k \, e^{-2\lambda_k \, q(T)}
  =0
  .
\label{eq:opt_qT}
\end{align}
%
%\begin{align}
%  m\bigl( q(T) \bigr)
%  +
%  \frac{2}{a_0 \, C_M \, M\bigl( q(T) \bigr)}
%  \sum_{k = 1}^{n-1}
%  \lambda_k \, \tilde u_k^2 \,
%  e^{-2 \, \lambda_k \, [a_0 \, T_0 + q(T)] }
%  \\=
%  \frac{1} { T \, (a_0 / 2) }
%  .
%\label{eq:opt_qT}
%\end{align}
%
\note{Let $q_T \equiv q(T)$,
$$
\begin{aligned}
  \frac{
    \partial \Err_R(T)
  }
  {
    \partial q_T
  }
  &=
  -a_0 \, M^2(q_T)
  -2
  \sum_{k=1}^{n-1} \lambda_k \,
  \tilde u_k^2 e^{-2 \, \lambda_k \, (a_0 \, T_0 + q_T)}
  ,
  \\
  \frac{
    \partial \Err_A(T)
  }
  {
    \partial q_T
  }
  &=
  \frac 2 T \,
  M(q_T) \,
  \int_0^{ q_T } M(q) \, dq
  =
  \frac{ 2 \, C_M } { T } \, M(q_T)
  .
\end{aligned}
$$
Then solve $\partial (\Err_R + \Err_A) / \partial Q_c = 0$.
}
%
We can rewrite the condition in terms of
the initial updating magnitude by Eq. \eqref{eq:mQ_invTa},
\begin{align}
  \alpha(0)
  =
  \frac{1}{T \, m\bigl( q(T) \bigr) }
  =
  \frac{ a_0 } { 2 }
  +
  \sum_{k = 1}^{n-1}
  \frac{
    \lambda_k \, \epsilon_k \, e^{-2\lambda_k \, q(T)}
  }{M^2\bigl( q(T) \bigr)}
  .
  \label{eq:opt_alpha0}
\end{align}
%
If the initial bias, $\epsilon_k$, is negligible,
which is typically true if no eigenvalue is close to $0$,
then
%
\begin{equation}
  \alpha( t = 0 )
  \approx
  \frac{ a_0 }
       { 2 }
  ,
\label{eq:half_alpha0}
\end{equation}
%
i.e. the optimal initial updating magnitude
is half of the equilibrium value.\footnote{This factor, $1/2$,
happens to coincide with the
recommended reduction factor of the updating magnitude
at stage transitions
in the original WL algorithm\cite{
wang2001, *wang2001pre}.
}

@@

The transition matrix method usually only for
simple sampling processes without major barrier
along the collective variable, $z$.
%
We may nonetheless estimate the magnitude of $\Gamma_k$
in a few special cases.
%
If sampling is perfect, $\Lambda_k = \delta_{k, 0}$,
then $\Gamma_k = 1$ for $k > 0$.
%
For a local sampling process,
if we borrow \eqref{eq:lambda_Gaussian}
for $\Lambda_k$,
with $\sigma_z$ interpreted as the average move size along $z$,
then
\begin{align}
\Gamma_k = \coth\left[
  \left(
    \frac{ \pi \, k \, \sigma_z } { g \, \Delta z}
  \right)^2
  \right].
\label{eq:Gamma_local}
\end{align}
For small $k$,
we have $\Gamma_k \propto \Delta z^2/(\sigma_z \, k)^2$,\cite{bussi2006}
which means that a doubled sampling range, $\Delta z$,
would make the histogram flattening
at least four times as difficult,
i.e. FDS methods are most effective
for short ranges\cite{wang2001, *wang2001pre}.
%
Using finer bins would, however, not change this value.
%
For a modestly large $k$,
$\Gamma_k \approx 1$.
and by Eq. \eqref{eq:xt2_eql},
we would expect a roughly flat tail
for the single-bin updating scheme
as $\langle \tilde v_k^2 \rangle \propto \Gamma_k$,
cf. Fig. \ref{fig:lj_xerr}.

\section{\label{sec:seplambda}
Updating scheme
with a set of well-separated eigenvalues}

We will study updating schemes
with a set of well-separated eigenvalues,
as a model for the large $k$ eigenmodes
of the Gaussian updating scheme.
%
We will show that the mass distribution will have
an inverse-$\bar q$ profile,
and derive a upper bound for the asymptotic error.

Starting from the definition, Eq. \eqref{eq:mass_func}, we have
$$
  M(\bar q)
  \le
  \sum_{k = 2}^n
  \sqrt{ \Gamma_k } \, \lambda_k \, e^{ -\lambda_k \, \bar q }
  .
$$
Thus,
it follows from Eq. \eqref{eq:mint} that, as $q(T) \to \infty$,
%
\begin{equation}
  C_M
  =
  \int_0^\infty
  M(\bar q) \, d \bar q
  \le
  \sum_{k = 1}^{n-1} \sqrt{ \Gamma_k }
  ,
  \label{eq:mint_ubound}
\end{equation}
%
and the asymptotic error from Eq. \eqref{eq:error_asym2}
is bounded as
\begin{equation}
  \Err_A(T)
  =
  \frac{ C_M^2 } { T }
  \le
  \frac 1 T
  \left(
    \sum_{k=1}^{n-1} \sqrt{ \Gamma_k }
  \right)^2
  .
  \label{eq:error_asym_ubound}
\end{equation}
%
It is readily seen that this bound is always
greater than the lower bound given by Eq. \eqref{eq:error_asym_singlebin},
%
$$
\frac{T}{(T+t_0)^2} \, \sum_{k=1}^{n-1} \Gamma_k
\le
\frac{1}{T} \sum_{k=1}^{n-1} \Gamma_k
\le
\frac{1}{T} \left( \sum_{k=1}^{n-1} \sqrt{ \Gamma_k } \right)^2
.
$$

Next, we will show that this bound represents
updating schemes with a set of very different eigenvalues.
%
First, we have for any mode $k^*$ $(k^* > 0)$,
%
\begin{align}
  M(\bar q)
  &
  \ge
  \sqrt{ \Gamma_{ k^* } } \,
  \lambda_{ k^* } \,
  e^{ -\lambda_{ k^* } \, \bar q }
  \ge
  \frac{ \sqrt{ \Gamma_{ k^* } } } { \bar q }
  ,
\label{eq:MQ_lbound1}
\end{align}
%
where the first inequality
would approach an equality if the $k^*$ mode
dominates the mass function, $M(\bar q)$,
and the second inequality
represent the limit of $\lambda_{k^*} \to 1/\bar q$.
%
Then, for the $k$th component of the asymptotic error,
we have from Eq. \eqref{eq:error_asym_Lagrangian}
%
\begin{align}
  \left\langle
  \tilde v_k^2(T)
  \right\rangle_A
  &=
  \int_0^T
  \Gamma_k \, \lambda_k^2 \, e^{2 \, \lambda_k \, [q(t) - q(T)] }
  \, {\dot q}^2(t) \, dt
  \notag
  \\
  &=
  \frac{ C_M } { T }
  \, \Gamma_k \, \lambda_k^2
  \int_0^{q(T)}
  \frac{ e^{-2 \, \lambda_k \, \bar q } } { M(\bar q) }
  \,  \, d \bar q
  \notag
  \\
  &\le
  \frac{ C_M }{ T } \,
  \frac{ \Gamma_k \, \lambda_k^2 } { \sqrt{ \Gamma_{k^*} } \, \lambda_{k^*} }
  \int_0^{q(T)}
    e^{-(2 \, \lambda_k - \lambda_{k^*} ) \, \bar q}
    \, d \bar q
  \notag
  \\
  &\le
  \frac{ C_M }{ T } \,
  \cdot
  \frac{ \Gamma_k }
  { \sqrt{ \Gamma_{k^*} } }
  \cdot
  \frac{ \lambda_k^2 }
  { \lambda_{k^*} \, (2 \, \lambda_k -\lambda_{k^*}) }
  ,
  \label{eq:y2_ubound_k}
\end{align}
%
where we have
used Eqs. \eqref{eq:Lagrangian_const} and \eqref{eq:MQ_lbound1}
on the second and third lines, respectively,
and extended the upper limit to infinity on the last line.
%
For a set of well-separated eigenvalues,
$\lambda_k$ depends on the mode index, $k$,
much more critically than $\Gamma_k$.
%
Thus, the tightest upper bound
can be approximately obtained from minimizing the last factor
on the right-hand side of Eq. \eqref{eq:y2_ubound_k},
and the minimal is reached at $k^* = k$,
when $\lambda_{k^*} = \lambda_k$.
%
So
%
\begin{equation}
  \left\langle
  \tilde v_k^2(T)
  \right\rangle_A
  \le
  %\approx
  \frac{ C_M }{ T }
  \sqrt{ \Gamma_k }
  .
  \label{eq:y2_ubound}
\end{equation}
%
By summing over $k$
and using Eq. \eqref{eq:mint_ubound},
we recover Eq. \eqref{eq:error_asym_ubound}.

From Eq. \eqref{eq:Gamma_local},
we know that $\Gamma_k$ would quickly approach a constant.
%
Then Eq. \eqref{eq:y2_ubound} shows that
the asymptotic error profile would also
approach a plateau in this case
(cf. Fig. \ref{fig:lj_xerr}
for a Gaussian example).
%
However, to reach this upper bound
the simulation has to be long enough
such that the upper limit of the integral
can be extended from $q(T)$ to infinity
in the last step of Eq. \eqref{eq:y2_ubound_k}.
%
As a result,
the flatter profile from Eq. \eqref{eq:y2_ubound}
would stop at mode $k$,
at which $\lambda_k \approx 1/q(T)$.

Finally, from Eqs. \eqref{eq:MQ_lbound1} and \eqref{eq:Gamma_local}
we know that for a large $\bar q$, which implies a sufficiently
large matching mode $k^*$ with $\lambda_{k^*} \approx 1/\bar q$,
the mass function would be roughly proportional to $1/\bar q$,
and so is the mass distribution, i.e.
\begin{equation}
  m(\bar q) \propto \frac{ 1} { \bar q },
  \qquad
  \mbox{for large $\bar q$.}
  \label{eq:m_invq}
\end{equation}
